Kids PM Summary

### 2020-05-28 child_health_data_cleaning.Rmd
Small plot and code annotation updates

---
title: "Kids health and child data cleaning"
author: "Ethan Walker"
date: "Started 27 Feb 2020, Updated 22 April 2020"
output: pdf_document 
---

@@

# Load and merge files from different locations
```{r}
wmt_athomelog_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_athomelog_20200227.xlsx")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "WMT",
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

@@

# Format at-home-log and merge with PM data
```{r}
kids_at_home_log <- 
  read_rds(paste0(file_path, "Output/kids_at_home_log.rds")) %>% 
  arrange(area, home_winter_id, at_home_day)

@@

@@

@@

@@

@@

at_home_pm_merged <- pm_clean %>% 
  group_by(area, home) %>% 
  mutate(pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE),
         pm_datetime_merge = round_date(pm_datetime_new, unit = "hour"),
         pm_obs_sampling_day = n()) %>% 
  separate(pm_datetime_merge, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home, sampling_day, sample_time) %>% 
  mutate(pm_mean_hour = mean(pm, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  distinct(sample_time, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(sampling_day = as.character(sampling_day)) %>% 
  select(area, home, home_winter_id, winter_id, treatment, sampling_day, 
         sample_date, sample_time, pm_mean_hour, pm_mean_sampling_day, 
         pm_mean_sampling_period, day_of_week, pm_obs_sampling_day, 
         pm_spike_1000, pm_spike_2000, pm_spike_5000) %>% 
  left_join(kids_at_home_log_new, 
            by = c("area", "home", "home_winter_id", "winter_id", "treatment",
                   "sampling_day", "sample_time")) %>% 
  group_by(area, home, home_winter_id, child_id_num, at_home) %>% 
  mutate(pm_at_home_sampling_period = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  group_by(area, home, home_winter_id, child_id_num, sampling_day, at_home) %>% 
  mutate(pm_at_home_sampling_day = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, child_id_num, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  distinct(at_home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  filter(!is.na(at_home)) %>% 
  arrange(area, home_winter_id, child_id_num, at_home, sampling_day)

@@

# Load and merge files from different locations
```{r}
wmt_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_medicalvisit_20200227.xlsx")) %>% 
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

>>>>

---
title: "Kids health and child data cleaning"
author: "Ethan Walker"
date: "Started 27 Feb 2020, Updated 28 May 2020"
output: pdf_document 
---

@@

# Load and merge files from different locations
```{r}
wmt_athomelog_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_athomelog_20200227.xlsx")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "WMT",
         at_home_day = as.factor(Day),
         # rename columns and recode as factor
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

@@

# Format at-home-log and merge with PM data
```{r}
# load data that was saved above
kids_at_home_log <- 
  read_rds(paste0(file_path, "Output/kids_at_home_log.rds")) %>% 
  arrange(area, home_winter_id, at_home_day)

@@

  # recode factors so 1=home, 0=not home; add overnight hours

@@

  # change data from wide to long, with row for hour of each day

@@

  # fill in overnight hours with assumption that child was home at night
  # match hour of day with corresponding location from the log

@@

  # recode overnight hours with assumption that child was at location they were
  # at 10pm and 7am

@@

# start with cleaned PM dataset
at_home_pm_merged <- pm_clean %>% 
  group_by(area, home) %>% 
  # average the PM for sampling period, day, and hour
  mutate(pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE),
         pm_datetime_merge = round_date(pm_datetime_new, unit = "hour"),
         pm_obs_sampling_day = n()) %>% 
  separate(pm_datetime_merge, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home, sampling_day, sample_time) %>% 
  mutate(pm_mean_hour = mean(pm, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  # keep one observation per hour of sampling, with PM averaged
  distinct(sample_time, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(sampling_day = as.character(sampling_day)) %>% 
  select(area, home, home_winter_id, winter_id, treatment, sampling_day, 
         sample_date, sample_time, pm_mean_hour, pm_mean_sampling_day, 
         pm_mean_sampling_period, day_of_week, pm_obs_sampling_day, 
         pm_spike_1000, pm_spike_2000, pm_spike_5000) %>% 
  # join with at home log from above
  left_join(kids_at_home_log_new, 
            by = c("area", "home", "home_winter_id", "winter_id", "treatment",
                   "sampling_day", "sample_time")) %>% 
  group_by(area, home, home_winter_id, child_id_num, at_home) %>% 
  # average PM during time child was at home
  mutate(pm_at_home_sampling_period = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  group_by(area, home, home_winter_id, child_id_num, sampling_day, at_home) %>% 
  mutate(pm_at_home_sampling_day = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, child_id_num, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  distinct(at_home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  filter(!is.na(at_home)) %>% 
  arrange(area, home_winter_id, child_id_num, at_home, sampling_day)

@@

# Load and merge files from different locations
```{r}
wmt_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_medicalvisit_20200227.xlsx")) %>% 
  # recode values to NA
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>%
  # recode temps of <90 to NA
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

### 2020-05-28 pm_summary.Rmd
Small plot and code annotation updates

model_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(area == "WMT") %>% 
  filter(home != "WMA300") %>% 
  distinct(home, .keep_all = TRUE)

@@

# Boxplots for PM by study area
```{r, fig.height=6, fig.width=8}
boxplots_kids_pm <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  ggplot(aes(area, pm)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
                 position=position_nudge(x = -0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
                 size=7, position=position_nudge(x = -0.35, y = 0)) +
    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
                 position=position_nudge(x = 0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
                 size=7, position=position_nudge(x = 0.35, y = 0)) +
    theme_minimal() +
    #ylim(0, 500) +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm
```

@@

```{r, fig.height=, fig.width=8}
boxplots_kids_pm <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  ggplot(aes(area, pm)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
    scale_y_log10() +
    theme_minimal() +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black",
                                      margin = margin(t = 0, r = 0, b = 30, l = 0)),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm

@@

>>>>

model_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(area == "WMT") %>% 
  filter(home == "WMA300") %>% 
  distinct(home, .keep_all = TRUE)

@@

# Boxplots for PM by study area
```{r, fig.height=6, fig.width=8}
boxplots_kids_pm <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(home != "WMA300") %>% 
  ggplot(aes(area, pm)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
                 position=position_nudge(x = -0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
                 size=7, position=position_nudge(x = -0.35, y = 0)) +
    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
                 position=position_nudge(x = 0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
                 size=7, position=position_nudge(x = 0.35, y = 0)) +
    theme_minimal() +
    #ylim(0, 500) +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm
```

@@

```{r, fig.height=, fig.width=8}
boxplots_kids_pm <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(home != "WMA300") %>% 
  ggplot(aes(area, pm_mean_sampling_period)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
    #scale_y_log10() +
    theme_minimal() +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black",
                                      margin = margin(t = 0, r = 0, b = 30, l = 0)),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm

@@

```{r}
# check PM spikes
spikes_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(home != "WMA300") %>% 
  select(area, home, sampling_day, pm_mean_sampling_day, pm_mean_sampling_period,
         pm_spike_1000, pm_spike_2000, pm_spike_5000,
         income_3level, stove_grade_3level, chimney_clean_3level, home_floors_2level) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  group_by(area, home) %>% 
  distinct(area, home, sampling_day, .keep_all = TRUE) %>% 
  mutate(pm_spike_1000 = if_else(pm_spike_1000 == "Yes", 1, 0),
         pm_spike_2000 = if_else(pm_spike_2000 == "Yes", 1, 0),
         pm_spike_5000 = if_else(pm_spike_5000 == "Yes", 1, 0),
         spike_sum_1000 = sum(pm_spike_1000),
         spike_sum_2000 = sum(pm_spike_2000),
         spike_sum_5000 = sum(pm_spike_5000)) %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  mutate(spike_sum_1000 = as.factor(spike_sum_1000),
         spike_sum_1000 = fct_collapse(spike_sum_1000, "0" = "0",
                                       "1" = "1",
                                       "2" = "2",
                                       "3+" = c("3", "5"))) %>% 
  ungroup() %>% 
  arrange(area, home)

spikes_summary <- spikes_data %>% 
  mutate(spike_sum_1000 = as.factor(spike_sum_1000),
         spike_sum_1000 = fct_collapse(spike_sum_1000, "0" = "0",
                                       "1" = "1",
                                       "2" = "2",
                                       "3+" = c("3", "5"))) %>% 
  group_by(spike_sum_1000) %>% 
  summarize("N" = n(),
            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE), 
            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE), 
            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
spikes_summary

table(spikes_data$spike_sum_1000)
table(spikes_data$spike_sum_1000, spikes_data$stove_grade_3level)
```
